---
title: "age_yoe_clustering"
author: "Christian Medeiros Adriano"
date: "May 4, 2017"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

### Cluster analysis
<p>Do we have meaningful clusters of Mechanical Turk workers that can be grouped by age and years of computer programming experience?</p>

```{r loadData, echo=FALSE}
setwd("C://Users//chris//OneDrive//Documentos//GitHub//programmerDemographics//WorkerCategorizationAnalysis");
data_all <- read.csv("demographics_data.csv",header = TRUE,sep=",");

data <- cbind(data_all$Years.Programming,data_all$Worker.Age);
colnames(data)<-c("years_of_programming_experience","age");
dataf = data.frame(data)
```
###Data cleaning
<p>Remove NA's, invalid age and years of experience, and age is smaller than years of experience</p>

```{r data, echo=FALSE}
#Remove rows with not available data
dataf <- dataf [rowSums(is.na(data))==0,]

#Remove Invalid age and invalid years of experience
dataf <- dataf [!dataf$years_of_programming_experience <0,]
dataf <- dataf [!dataf$age <0,]

#Remove people with no experience in programming
dataf <- dataf [!dataf$years_of_programming_experience ==0,]

#Remove people who whose age and experience 
dataf <- dataf [!(dataf$age - dataf$years_of_programming_experience) <0,]

#Remove outliers
dataf <- dataf [!dataf$years_of_programming_experience >40,]
dataf <- dataf [!dataf$age >80,]

summary(dataf);
cat("number of workers:",length(dataf[,1]));

```

###Plotting and regression curve
```{r plotting, echo=FALSE}
ggplot(dataf, aes(x=years_of_programming_experience, y=age)) +
  geom_point(shape=1) +    # Use hollow circles
  geom_smooth() +            # Add a loess smoothed fit curve with confidence region
  ggtitle("Distribution of age by years of experience")
```
<p>Many people seemed to have reported their experiences in multiples of 5 years.</p>


## Clustering
```{r plotKMeansClusters, echo=FALSE}
#Function computes N clusters and plot a chart with them

plotKMeansClusters<- function(numberOfClusters){
  (cl <- kmeans(dataf,numberOfClusters))
  
  dataf$cluster=factor(cl$cluster)
  centers=as.data.frame(cl$centers)
  
  #Looking at the clusters with size diameter 15.
  ggplot(dataf, aes(x=years_of_programming_experience, y=age, color=cluster)) +
    geom_point(shape=1) +    # Use hollow circles
    geom_smooth() + # Add a loess smoothed fit curve with confidence region
    geom_point(data=centers, aes(x=years_of_programming_experience,y=age, color='Center'), size=40, alpha=.3)+
    ggtitle("Clusters of years of experience and age")+
    guides(color=FALSE);
  
}
```

<p>What happens if we divide workers in 5 clusters, which is the number of professions?</p>
```{r plotKMeansClusters5, echo=FALSE}
plotKMeansClusters(5)
```
<p>We can see a lot of superposition of clusters. So, we let's try fewer clusters</p>
```{r plotKMeansClusters432, echo=FALSE}
plotKMeansClusters(4)
plotKMeansClusters(3)
plotKMeansClusters(2)
```

<p>
Only with two clusters we circles do not superpose. Interpreting that, we would have two large groups of workers. People above 35 years old with a wide spread of programming experience. While people below 35 concentrated from 1 to 15 years of programming experience.
</p>

####Future analysis
<ul>
<li>Do these clusters map to professions? In other words, can we predict the worker profession based on age and years of programming experience?</li>
<li>Would it make much difference in terms of number of meaningful clusters if I apply SVM or Decision trees methods?</li>
<li>Are worker clusters statistically significanlty distinct in terms of average accuracy?</li>
</ul>

<hr width="50%">
<p>Thanks for reading! Is there anything I could improve? Please leave a comment - Christian.</p>
<i>"We are trying to prove ourselves wrong as quickly as possible, because only in that way we can find progress." Richard P. Feynman</i>
